---
title: "tidy-model"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('tidyverse') ; library('modeldata')
``` 

## Introduction 

This R Markdown document demonstrates my abilities to use models for data analysis using a data set collected on crickets

```{r load data, include = FALSE}
data(crickets, package = "modeldata")
names(crickets)
```

## What is that Sound?

In this report, we examine what factors predict a cricket's chirp rate

```{r summary, echo = FALSE}
summary(crickets)
ggplot(crickets, aes(x = rate)) +
  geom_histogram(bins = 8) + 
  ggtitle("Distribution of the chirp rate of crickets") +
  xlab('Chirp rate (per min.)')
```

## Temperature affects chirp rate

```{r temp, echo= FALSE}
ggplot(crickets, aes(x = temp, y = rate)) +
  geom_point() + 
  geom_smooth(method = 'lm') +
  ggtitle("Plot of temperature and chirp rate") +
  ylab('Chirp rate (per min.)') +
  xlab('Temperature (Celsius)')

temp_lm <- lm(rate ~ temp, crickets)

summary.lm(temp_lm)
```

Based on a scatter plot of temperature and chirping and a correlation test, it seems that as temperature increases one degree, the rate of chirping increases about 4.2 chirps per minute.

## Species-specific effects of temperature on chirping

```{r species, echo= FALSE}
ggplot(crickets, aes(x = temp, y = rate, color = species)) +
  geom_point() + 
  geom_smooth(method = 'lm') +
  ggtitle("Plot of temperature and chirp rate for two species of crickets") +
  ylab('Chirp rate (per min.)') +
  xlab('Temperature (Celsius)')
  
  
species_lm <- lm(rate ~ temp + species, crickets)

summary.lm(species_lm)
```

The rate of chirping increased based on temperature. However, the O.exclamatonis seems to have a higher chirping rate as a species compared to the other

## Interactions

```{r species historgram, echo = FALSE}
ggplot(crickets, aes(x = rate, fill = species)) +
  geom_histogram(position = 'identity', alpha = 0.7, bins = 8) + 
  ggtitle("Distribution of the chirp rate of crickets") +
  xlab('Chirp rate (per min.)')
```

It shows that the O.exlamatonis has more counts of their species in counts that have a higher rate

```{r interactions, echo= FALSE}
ggplot(crickets, aes(x = temp, y = rate, color = species)) +
  geom_point() + 
  geom_smooth(method = 'lm') +
  ggtitle("Plot of temperature and chirp rate for two species of crickets") +
  ylab('Chirp rate (per min.)') +
  xlab('Temperature (Celsius)')
  
species_x_temp_lm <- lm(rate ~ temp + species + temp:species, crickets) 
summary.lm(species_x_temp_lm)

anova(species_lm, species_x_temp_lm)
```